---
title: "Dynamic Documents - R Markdown"
subtitle: "Part 3 - Notebooks using R - Practice"
author: "Fabien Forge"
date: "15/10/2021"
output: html_document

knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "./../../Presentations/Day 2 - Dynamic Documents")
  })
---

# Practice

- Let's use a (slightly amended) dataset from Nathan Nunn's [website](https://scholar.harvard.edu/nunn/pages/data-0)

- We will use the data from is famous 2008 QJE paper: "The Long-Term Effects of Africa's Slave Trades"


## Some libraries  
```{r, message=F}
library(tidyverse)
library(ggplot2)
library(skimr)
library(jtools)
library(huxtable)
library(sandwich)
```


## Read the data in memory

```{r, results='asis', message=FALSE, out.width="50%"}
path.url <- "https://www.dropbox.com/s/oluf7cdidu2hq2g/nunn_slave_trade.csv?dl=1"

df <- read.csv(path.url)

head(df)
```

## Select columns and Pipe Operator

```{r}
# select country, gdppercapita, slaveexports, coastline, popdensity, legor_uk, legor_fr

df <- df %>% select(country, gdppercapita, slaveexports, coastline, popdensity, legor_uk, legor_fr)
head(df)

```

## Mutate 
```{r}
# take log of gdppercapita

df <- df %>% mutate(ln_gdgpc=log(gdppercapita))
head(df)
```

## Groupby and Summarize 

```{r}
# average gpd per capita per legal origin
df %>% group_by(legor_uk) %>% 
  summarise(mean_gdp=mean(gdppercapita))

```


# A simple regression

- We will first consider the effect of historical slave exports on countries' per capita GDP in 2000 using a simple linear regression:
<br><br>
$Y_i = \beta_0 + \beta_1 X_i + u_i$
<br><br>
$\text{GDP}_i = \beta_0 + \beta_1 \text{SlaveExport}_i + u_i$ 

```{r }
# regress gdppercapita on slaveexports
results <- lm("gdppercapita ~ slaveexports", data=df)
summary(results)
```



# Heteroskedastic-Robust Standard Errors
```{r }
results <- lm("gdppercapita ~ slaveexports", data=df)
summ(results, robust = 'HC1')
```


# An additional regressor


- Might the length of a country's coastline (normalized by its area) be an omitted variable here?
  - why?

```{r}

results2 <- lm( "gdppercapita ~ slaveexports + coastline", data=df)
summ(results2, robust = 'HC1')
```




# A third regressor
```{r}
results3 <- lm( "gdppercapita ~ slaveexports + coastline+ popdensity", data=df )
summ(results3, robust = 'HC1')
results4 <- lm( "gdppercapita ~ slaveexports + coastline+ popdensity + legor_uk", data=df )
summ(results4, robust = 'HC1')
```





```{r}
export_summs(results, results3, model.names = c("Simple Regression", "Multiple Regression"), robust='HC1')
```



# Creating regression tables
- There are many solutions in R to create regression tables
<br><br>
- Today we will use the export_summ() function built on huxtable and jtools packages

# Gather model estimations

- Say that you want to present our first vs our third regression

```{r }

```


# Model names

```{r }

```


# Stars
- Let's use the thresholds traditionally used in development

```{r }


```



# Do not report the intercept
```{r }


```


# Let's add the F-statistic

- And the associated p-value

```{r }

```


# Add a third model

```{r }


```


